On Adding Pattern Matching to Haskell-based Deeply Embedded Domain Specific Languages\\ \large{Not For Distribution}

### 2020-07-25 paper/paper.tex
First cut at control flow diagram
Restructuring description of plugin

\clearpage
\subsection{Overall Control Flow of the Core Plugin}

In the following flowchart, the names in parentheses in a node refers to the names of transformations from the following section (Section 8.2). Note that
calls to \ttt{unrep :: E a -> a} are only used internally and will not exist in the final result of the transformation.

\begin{tikzpicture}[node distance=2cm]
\node (start) [startstop] {Core from GHC};
  \node (tailrec) [process, below of=start, yshift=-0.5cm]{($D, T_{\ttt{f}}, T'_{\ttt{f}}$) Transform tail recursion of definitions that have a mark (when applicable)};
  \node (findmark) [process, below of=tailrec, yshift=-0.5cm]{($C^*$) Find marked expression and erase mark};
    \node (handlecase) [process, below of=findmark]{($C, M$) Mark scrutinee};
    \node (handleprim) [process, right of=handlecase, xshift=1.6cm]{($C$) Encode as EDSL primitive};
    \node (handleconstr) [process, right of=handleprim, xshift=1.6cm]{($C$) Wrap with \ttt{construct}};

    \node (marksubs1) [process, below of=handleprim]{($C, M$) Mark sub-expressions};
    \node (marksubs2) [process, below of=handleconstr]{($C, M$) Mark sub-expressions and wrap with \ttt{unrep}};
    \node (betareduce) [process, below of=marksubs2]{Beta reduce \ttt{(construct ...)} expression and simplify};
    \node (repunrep) [process, below of=betareduce]{Replace expressions of the form \ttt{(rep (unrep x))} with \ttt{x}};

    \node (onebranch) [process, below of=handlecase]{($B$) Transform one branch};
    \node (prodmatch) [process, below of=onebranch]{($P$) Transform one field in product match of current branch};
    \node (abstractvar) [process, below of=prodmatch, yshift=-0.5cm]{($P, A$) Abstract the pattern variable of the field with \ttt{Lam} and \ttt{Var}};
    \node (markbody) [process, below of=abstractvar]{($P, M$) Mark branch body};
    \node (endstep) [process, below of=markbody, right of=markbody, xshift=1.6cm]{End of one transformation step};

  \draw [arrow] (start) -- (tailrec);
  \draw [arrow] (tailrec) -- (findmark);
  \draw [arrow] (findmark) -- node[anchor=east] {Primitive} (handleprim);
  \draw [arrow] (findmark) -- node[anchor=west, yshift=0.1cm] {Constructor/constructor application} (handleconstr);
  \draw [arrow] (findmark) -- node[anchor=east] {\ttt{case}} (handlecase);

  \draw [arrow] (handleprim) -- (marksubs1);

  \draw [arrow] (handleconstr) -- (marksubs2);

  \draw [arrow] (handlecase) -- (onebranch);
  \draw [arrow] (onebranch) -- (prodmatch);
  \draw [arrow] (prodmatch) -- (abstractvar);
  \draw [arrow] (abstractvar) -- (markbody);

  \draw [arrow] (marksubs2) -- (betareduce);
  \draw [arrow] (betareduce) -- (repunrep);

  \draw [arrow] (marksubs1) -- (endstep);
  \draw [arrow] (repunrep) -- (endstep);
  \draw [arrow] (markbody) -- (endstep);

  % \draw (prodmatch) edge[bend right, left] node {While fields remain} (abstractvar);
  % \draw [arrow] (abstractvar.east) to ++(0.5,0) |- node[anchor=west] {While fields remain} (prodmatch.east);
  \draw [arrow] (abstractvar.west) -| ++(-1.5,0) |- node[anchor=south] {While fields remain} (prodmatch.west);

  \draw [arrow] (markbody.west) -| ++(-3,0) |- node[anchor=east, anchor=south] {While branches remain} (onebranch.west);

  \draw [arrow] (endstep.west) -| ++(-8.5,0) |- node[anchor=west, anchor=south] {While marked expressions remain} (findmark.west);

  % \draw (findmark)  edge[loop left] node {While marked expressions remain} (findmark);

  % \draw (findmark)  edge[loop left] node {While marked expressions remain} (findmark);
  % \draw (onebranch) edge[loop left] node {While branches remain} (onebranch);
  % \Loop[label=\(While\, fields\, remain\), labelstyle={left=5pt}](prodmatch);
  % \path (bodylam) edge [left] (onebranch);
  % \path (bodylam) edge [loop above] node {test} (onebranch);
  % \Loop[label=\(While\, remaining\, branches\, exist\), labelstyle={left=5pt}](onebranch)
  % \draw [arrow] (bodylam) -- (onebranch);
  % \draw [arrow] (bodylam.east) to  ++ (1.5,0) |- node[anchor=south] {While branches remain} (onebranch);
  % \draw [arrow] (bodylam.east) to  ++ (1.5,0) |- node[anchor=south] {While branches remain} (onebranch);
  % \Loop[label=\(While\, branches\, remain\), labelstyle={left=5pt}](onebranch);
  % \Loop[label=\(While\, marked\, expressions\, remain\), labelstyle={left=5pt}](findmark);
\end{tikzpicture}

\clearpage
\subsection{Syntactic Transformations of the Core Plugin}

\begin{align*}
  D&\expr{\ttt{f x = internalize (externalize (case s of \{ ... \}))}}\\
      &\rewrites \begin{cases}
        \ttt{f = internalize $C^{*}\expr{\ttt{externalize (runIter ($\lambda \ttt{x} \rarr \ttt{case s of \{ $T'_{\ttt{f}}\exprdots$}$ \}))}}$} & \text{if \ttt{f} occurs free in \ttt{\{ ... \}}}\\
        \ttt{f x = internalize $C^{*}\expr{\ttt{externalize (case s of \{ ... \})}}$} & \text{otherwise}
      \end{cases}
\end{align*}

@@

>>>>

In the flowchart given in Figure~\ref{fig:CorePlugin}, the names in parentheses in a node refers
to the names of transformations given in the rewrite rules in Figure~\ref{fig:Rewrites}. Calls
to \ttt{unrep :: E a -> a} are only used internally and will not exist in
the final result of the transformation.

@@

\clearpage
\begin{figure*}
  \centering 
  \resizebox{2.0\columnwidth}{10cm}{\includegraphics{Diagrams/Flowchart.pdf}}
   \caption{Control Flow of the Core Plugin}
   \label{fig:CorePlugin}
\end{figure*}

% \subsection{Syntactic Transformations of the Core Plugin}

\begin{figure}[hp]
  \centering
\begin{align*}
  D&\expr{\ttt{f x = internalize (externalize (case s of \{ ... \}))}}\\
      &\rewrites \begin{cases}
        \ttt{f = internalize $C^{*}\expr{\ttt{externalize (runIter ($\lambda \ttt{x} \rarr \ttt{case s of \{ $T'_{\ttt{f}}\exprdots$}$ \}))}}$} & \text{if \ttt{f} occurs free in \ttt{\{ ... \}}}\\
        \ttt{f x = internalize $C^{*}\expr{\ttt{externalize (case s of \{ ... \})}}$} & \text{otherwise}
      \end{cases}
\end{align*}

\captionsetup{justification=centering}
\caption{Rewrite rules}
\label{fig:Rewrites}
\end{figure}

% \clearpage
